---
title: "GitHub R Intro"
author: "Jordyn Geller"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
source("setup.R")
```

```{r}
data("penguins")
penguins
```

```{r}
class(penguins)
str(penguins)
```

```{r}
class(penguins$species)
str(penguins$species)
```

```{r}
myList <- list("apple", 1993, FALSE, penguins)
str(myList)
```

```{r}
list(myList, list("more stuff here", list("and more")))
```

```{r}
names(myList) <- c("fruit", "year", "logic", "data")
names(myList)

```
# Indexing

# for lists we use double brackes [[]]
```{r}
myList[[1]]

myList[["data"]]
```

# for vectors we use single brackets []
```{r}
myVector <- c("apple", "banana", "pear")
myVector[2]
```
# index data frames

# dataframe[row(s), columns()]
```{r}
penguins[1:5, 2]

penguins[1:5, "island"]

penguins[1, 1:5]

penguins[1:5, c("species","sex")]

penguins[penguins$sex=='female',]
```

# $ for a single column
```{r}
penguins$species
```

# 6.3.1 Exercises

# 1) Why don’t the following lines of code work? Tweak each one so the code runs.
```{r}
myList[["fruit"]]
```
```{r}
penguins$flipper_length_mm
```

```{r}
penguins[penguins$island=='Dream',]
```

# 2) How many species are in the penguins dataset? What islands were the data collected for? (Note: the unique() function might help)
```{r}
length(unique(penguins$species))
```
# 3) Use indexing to create a new data frame that has only 3 columns: species, island and flipper length columns, and subset all rows for just the ‘Dream’ island.
```{r}

```

# dataframe[row(s), columns()]
```{r}
penguins[penguins$island =="Dream", c("species","island", "flipper_length_mm")]
```

# 4) Use indexing and the mean() function to find the average flipper length for the Adelie species on Dream island. (Note: explore the mean() function and how to deal with NA values).
```{r}
avgflipper <- penguins[penguins$island == "Dream" & penguins$species == "Adelie",]
mean(avgflipper$flipper_length_mm)
```

# 6.4 The dplyr package

```{r}
library(dplyr)
```

# Subset rows with filter:
```{r}
filter(penguins, species == "Chinstrap")

filter(penguins, species != "Chinstrap")

filter(penguins, island %in% c("Dream", "Biscoe") & !is.na(bill_length_mm))
```

# Select columns with select():
```{r}
?select
```

# Select two specific variables
```{r}
select(penguins, species, sex)
```

# Select a range of variables
```{r}
select(penguins, species:bill_length_mm)
```

# Rename columns within select
```{r}
select(penguins, gender = sex, island)
```

# Select column variables that are recorded in mm a
```{r}
select(penguins, contains("mm"))
```


# Create new variables with mutate():

# New variable that calculates bill length in cm
```{r}
mutate(penguins, bill_length_cm = bill_length_mm/10)
```
# New variable that calculates bill length in km
```{r}
mutate(penguins, bill_length_km = bill_length_mm/1000000)
```

# mutate based on conditional statements paste0()
```{r}
mutate(penguins, species_sex = if_else(sex == 'male', paste0(species,"_m"), paste0(species, "_f", sep="-")))
```

# mutate based on conditional statements paste()
```{r}
mutate(penguins, species_sex = if_else(sex == 'male', paste(species,"_m"), paste(species, "_f", sep="-")))
```

# 6.4.1 The pipe %>%

```{r}
df1 <- filter(penguins, island == "Dream")
df2 <- mutate(df1, flipper_length_cm = flipper_length_mm/10)
df3 <- select(df2, species, year, flipper_length_cm)
print(df3)
```

```{r}
penguins %>% 
  filter(island == "Dream") %>% 
  mutate(flipper_length_cm = flipper_length_mm/10) %>% 
  select(species, year, flipper_length_cm)
```

# group_by(species)
```{r}
penguins %>% 
  group_by(species) %>% 
  summarise(body_mass_avg = mean(body_mass_g, na.rm = TRUE))
```

# group_by(islands)
```{r}
penguins %>% 
  group_by(island) %>% 
  summarise(body_mass_avg = mean(body_mass_g, na.rm = TRUE))
```

# group_by(species) - count
```{r}
penguins %>% 
  group_by(species, year) %>% 
  summarise(n_observations = n())
```


# 6.3.2 Exercises

# 1) Reorder the variables in penguins so that year is the first column followed by the rest (Hint: look into the use of everything()).
```{r}
?everything
```

```{r}
penguins %>%
  select(year, everything())
```

# 2) Create a new column called ‘size_group’ where individuals with body mass greater than the overall average are called ‘large’ and those smaller are called ‘small’.
```{r}
penguins %>%
  mutate(size_group = if_else(body_mass_g > mean(body_mass_g, na.rm = TRUE), "large", "small"))
```

# first try
penguins %>%
  summarise(body_mass_avg = mean(body_mass_g)) %>%
  mutate(size_group, body_mass_g > body_mass_avg, paste(body_mass_g, "large")) %>%
  mutate(size_group, body_mass_g < body_mass_avg, paste(body_mass_g, "small"))


# 3) Find out which year for each species were individuals on average the largest according to body mass.
```{r}
penguins %>%
  group_by(species) %>%
  filter(body_mass_g == max(body_mass_g, na.rm = TRUE))
```


# 4) You want to filter data for years that are not in a vector of given years, but this code doesn’t work. Tweak it so that it does. (Yes, you could just filter year to equal 2007 in this case but there is a trouble-shooting lessons here).

```{r}
penguins %>% 
  filter(year %in% c(2007))
```


# 7. Visualize

```{r}
ggplot(penguins) + 
  geom_histogram(mapping = aes(x = flipper_length_mm))
```
# Histogram example: flipper length by species
```{r}
ggplot(penguins) +
  geom_histogram(aes(x = flipper_length_mm, fill = species), alpha = 0.5, position = "identity") +
  scale_fill_manual(values = c("darkorange","darkorchid","cyan4"))
```
# create a seperate plot for each species
```{r}
ggplot(penguins) +
  geom_histogram(aes(x = flipper_length_mm, fill = species), alpha = 0.5, position = "identity") +
  scale_fill_manual(values = c("darkorange","darkorchid","cyan4")) +
  facet_wrap(~species)
```
```{r}
ggplot(penguins) +
  geom_bar(mapping = aes(x = island, fill = species))
```

```{r}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"), 
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

```{r}
ggplot(penguins) +
  geom_point(mapping = aes(x = body_mass_g, y = flipper_length_mm, color = species))
```


# 7.1 Exercises

# 1) Make a barplot showing the average flipper length for each species.
```{r}
ggplot(penguins) +
 geom_bar(mapping = aes(x = species, fill =  mean(flipper_length_mm)))
```


# 2) Make a scatter plot of bill length compared to bill depth but only for observations on the Dream island.
```{r}
ggplot(penguins) +
  geom_point(mapping = aes(x = bill_length_mm, y = bill_depth_mm, fill = 'Dream'))
```

```{r}
penguins %>%
  filter(island == "Dream") %>%
  ggplot() +
  geom_point(mapping = aes(x = bill_length_mm, y = bill_depth_mm))
```


