---
title: "GitHub R Intro"
author: "Jordyn Geller"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
source("setup.R")
```

```{r}
data("penguins")
penguins
```
```{r}
class(penguins)
str(penguins)
```
```{r}
class(penguins$species)
str(penguins$species)
```
```{r}
myList <- list("apple", 1993, FALSE, penguins)
str(myList)
```
```{r}
list(myList, list("more stuff here", list("and more")))
```
```{r}
names(myList) <- c("fruit", "year", "logic", "data")
names(myList)

```
# Indexing

# for lists we use double brackes [[]]
```{r}
myList[[1]]

myList[["data"]]
```

# for vectors we use single brackets []
```{r}
myVector <- c("apple", "banana", "pear")
myVector[2]
```
# index data frames

# dataframe[row(s), columns()]
```{r}
penguins[1:5, 2]

penguins[1:5, "island"]

penguins[1, 1:5]

penguins[1:5, c("species","sex")]

penguins[penguins$sex=='female',]
```

# $ for a single column
```{r}
penguins$species
```

# 6.3.1 Exercises

# 1. Why don’t the following lines of code work? Tweak each one so the code runs.
```{r}
myList[["fruit"]]
```
```{r}
penguins$flipper_length_mm
```

```{r}
penguins[penguins$island=='Dream',]
```

# 2. How many species are in the penguins dataset? What islands were the data collected for? (Note: the unique() function might help)
```{r}
length(unique(penguins$species))
```
# 3. Use indexing to create a new data frame that has only 3 columns: species, island and flipper length columns, and subset all rows for just the ‘Dream’ island.
```{r}

```

# dataframe[row(s), columns()]
```{r}
penguins[penguins$island =="Dream", c("species","island", "flipper_length_mm")]
```

# 4. Use indexing and the mean() function to find the average flipper length for the Adelie species on Dream island. (Note: explore the mean() function and how to deal with NA values).
```{r}
avgflipper <- penguins[penguins$island == "Dream" & penguins$species == "Adelie",]
mean(avgflipper$flipper_length_mm)
```

# 6.4 The dplyr package

```{r}
library(dplyr)
```

# Subset rows with filter
```{r}
filter(penguins, species == "Adelie")

filter(penguins, species != "Adelie")

filter(penguins, island %in% c("Dream", "Torgersen") & !is.na(bill_length_mm))
```

# Subset columns with select()
```{r}
?select
```

```{r}

```


